---
title: 'Data Products: Plotly Assignment'
author: "Noel Temena"
date: "September 9, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Plotly
Machine Learning sample output taken from a dummy data set.
Graph shown are a visual representation of the different accuracy values.

```{r  echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)

letter_fact <- factor(LETTERS[1:5]) # dummy factor observation
n <- 500 # default number of rows or observation
df <- data.frame(Category = replicate(n,sample(letter_fact,1))) # setup a data frame with rows and values
df$y <- df$Category # create a 2nd column using the same values form the 1st column

change_col_y <- function(value)
{
   if (value == "A") {answer <- sample(c("B","E"),1,prob = c(0.9, 0.1))}
   else if (value == "B") {answer <- sample(c("C","D","E"),1,prob = c(0.4, 0.4,0.2))}
   else if (value == "C") {answer <- sample(c("B","D"),1,prob = c(0.8, 0.2))}
   else if (value == "D") {answer <- sample(c("C","E"),1,prob = c(0.9, 0.1))}
   else if (value == "E") {answer <- sample(c("D","A"),1,prob = c(0.9, 0.1))}
   returnValue(answer)
}

plot_data <- function(number)
{   
   plot_title <- paste("Machine Learning: ", number, "% predictive analysis accuracy")
   p <- qplot(x= df$y,y= df$Category, data = df, xlab = "Predicted", ylab = "Actual", geom = c("boxplot", "jitter"), colour= Category)
   p <- ggplotly(p)
   p
}
# main= plot_title,

get_pointer <- function(num1){
   df$y <- df$Category
   result <- sample(1:500,((100 - num1)  / 100) * 500)
   returnValue(result)
}
```

## Ideal Plot 1
Sample Machine learning output with 100% predictive accuracy 
```{r  echo=FALSE, message=FALSE, warning=FALSE}
plot_data(100)
```



## Plot 2
 Sample Machine learning output with 96% predictive accuracy 
```{r  echo=FALSE, message=FALSE, warning=FALSE}
y_pointer <- get_pointer(96)
df$y <- df$Category
df$y[y_pointer] -> y_values 
df$y[y_pointer] <- sapply(y_values, change_col_y)
plot_data(96)
```   

## Plot 3
 Sample Machine learning output with 77% predictive accuracy 
```{r  echo=FALSE, message=FALSE, warning=FALSE}
y_pointer <- get_pointer(77)
df$y <- df$Category
df$y[y_pointer] -> y_values 
df$y[y_pointer] <- sapply(y_values, change_col_y)
plot_data(77)
```

## Plot 4
 Sample Machine learning output with 60% predictive accuracy 
```{r  echo=FALSE, message=FALSE, warning=FALSE}
y_pointer <- get_pointer(60)
df$y <- df$Category
df$y[y_pointer] -> y_values 
df$y[y_pointer] <- sapply(y_values, change_col_y)
plot_data(60)
```